---
title: "Hotel Booking Prediction Case Study"
author: "LJ"
date: "2023-02-25"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      # cache = TRUE,
                      autodep = TRUE,
                      # cache.lazy = FALSE,
                      message = FALSE, 
                      include = TRUE,
                      dpi = 180,
                      fig.width = 8, 
                      fig.height = 5)

pacman::p_load(tidyverse,       # for all the tidymodels packages
               janitor,         # for all cleaning data steps
               here,            # for relative directory
               readxl,          # to read in excel files
               kableExtra,      # to make beautiful tables
               tidymodels,      # for all the tidymodels packages
               skimr,           # to get summary data
               vip)             # to get variable importance data

ggplot2::theme_set(theme_minimal())

```

## Introduction

This notebook will reproduce the case study available in the
`tidymodels` website, called ["A predictive modeling case
study"](https://www.tidymodels.org/start/case-study/) using a [version
of the data
set](https://gist.github.com/topepo/05a74916c343e57a71c51d6bc32a21ce)
called Hotel Booking from a study conducted by [Antonio, Almeida, and
Nunes (2019)](https://doi.org/10.1016/j.dib.2018.11.126)

In this study, all of the steps for creating a predictive model are
used. This section is the last of the [Getting
Started](https://www.tidymodels.org/start/) part of the `tidymodels`
website. The previous sections covered one step of the predictive model
creation separately and in detail.

I will attempt to follow the example here since I have a hard time not
seeing the whole process holistically.

## Loading the data

To read the data into R we will use the `readr::read_csv()` to creat a
copy in my repository.

```{r reding the hotel data into R as csv}
# read the hotel data
hotels <- 
  read_csv("https://tidymodels.org/start/case-study/hotels.csv") %>% 
  mutate(across(where(is.character), as.factor))

# save a copy of the dataset as RDS
write_csv(hotels, path = "data/hotels.csv")

```

```{r}
# save a copy of the dataset as RDS
write_csv(hotels, file = "data/hotels.csv")
```

## Looking at the date

```{r check dimensions}
# Check the dimensions of the resulting data set
hotels %>% 
  dim() %>% 
  as_tibble() %>% 
  add_column(element = c("rows", "columns"),
             .before = "value")

```

The resulting dataset has 50 000 rows and 23 columns. The modification
of the original consisted of removing cancelled reservations only to
concentrate on hotel stays that included children.

```{r looking into the dataset}
# Looking into the dataset
hotels %>% 
  dplyr::glimpse()

```

```{r looking into the data using skimr}
# Looking into the data using skimr
hotels %>% 
  skimr::skim()

```

The main objective is to predict which hotel actual stays included
children or babies and which did not.

Let's check the variable children.

```{r checking the variable children}
hotels %>% 
  janitor::tabyl(children) %>% 
  janitor::adorn_percentages("col", na.rm = TRUE, percent) %>% 
  janitor::adorn_pct_formatting(,,,percent) %>% 
  dplyr::as_tibble()

```

The dataset show an imbalance since only 8.1% of our date are hotel
stays that contain children. This can be tackled with functions within
`tidymodels` such as `recipes` however, this will be used as-is.

Once this has been done then we're ready to start creating our models.\
As seen in the first session where we followed [Rebecca Barter's
example](02_tidymodels-tidy-machine-learning.html), the steps to follow
will be:

-   Getting our train, validation, and testing datasets

-   Resampling

-   Define the recipe

-   Specifying the model

-   Creating a workflow

-   Tune the model

-   Evaluating our resulting model

-   Using the model

## Data Splitting and Resampling

WIP
